%===============================================================
\chapter{Academic Entrepreneurial Ventures}
%===============================================================

\begin{chapterabstract}
	This chapter motivates the thesis by examining the roles and impacts of start-ups and spin-outs emerging from academic environments as powerful vehicles for transforming innovative ideas and research into market-ready products and services. In academia, these ventures not only drive technological and economic progress but also provide students with hands-on entrepreneurial experience and alternative career pathways. However, the process of launching and sustaining such ventures is shaped by national innovation ecosystems, regulatory environments, and varying institutional support systems. The chapter also analyses the challenges and opportunities faced by ventures in the Czech Republic’s innovation landscape and the game development sector.
\end{chapterabstract}

As universities continue to expand their role beyond traditional teaching and research, there is a growing emphasis on their capacity to drive innovation and contribute to local economic development. One of the broader aspirations of this work is to strengthen the local entrepreneurial ecosystem by empowering students to transform creative ideas into viable ventures. In this context, understanding these mechanisms and their potential benefits becomes essential.

%===============================================================
% \section{Start-Ups and Spin-Outs}
\textbf{Start-up} is a new business aiming to develop a unique product or service to meet market needs. Characterized by innovation, high growth potential and often low revenue, they typically operate in uncertain environments and rely on venture capital\footnote{Venture capital is private equity investment into a project with a substantial element of risk, that is believed to have growth potential.}, angel investors\footnote{Angel investors are willing to back early stage start-ups when other investors aren't.}, or other funding sources to support their development.\cite{investopedia_start-ups}

\textbf{Spin-out} is a new company created from an existing organisation, such as a university or a corporation, formed to commercialise developed research, intellectual property, or technology. Spin-outs benefit from the resources and expertise of their parent organisations while functioning as independent entities.\cite{mfmac_spin-out}

In academia, start-ups and spin-outs translate ideas and research into marketable products and services. Universities and research institutions support these ventures through incubators, technology transfer offices, and funding programmes. Academic spin-outs, in particular, utilise faculty expertise, patents, and support to drive innovation and economic impact.

%===============================================================
\section{Impact of Academia-Driven Entrepreneurship}
Student spin-outs transfer theoretical early-stage research from universities to real-world applications. They are seen as an attractive alternative to patent licensing as they are more likely to impact the local economy. Spin-outs create jobs for highly skilled workers and provide valuable knowledge spillover for other companies.\cite{Caputo_spin-offs-impact}

Students taking part in spin-outs and start-ups can gain experience in entrepreneurship, practise business planning, market analysis, technology commercialisation and more. Even unsuccessful undertakings provide valuable learning opportunities. Moreover, venturing offers an additional career path for students (particularly researchers facing limited job opportunities).\cite{kauffman_tech-transfer}

Academic ventures foster collaboration between students, faculty, and external actors such as industry partners. This strengthens the university's ecosystem for innovation and entrepreneurship. Successful undertakings improve the university's reputation and potentially attract more funding for research and other programmes.

Academic ventures have demonstrated the ability to adapt quickly to societal challenges. For example, during the COVID-19 pandemic, some quickly developed solutions to address urgent problems. Many also focus on solving pressing global problems, such as sustainability or healthcare challenges, contributing to broader societal benefits.\cite{Battaglia_spin-off-response}

Commercial undertakings of students and academic staff offer many benefits, from driving local economic growth to fostering innovation and creating new career opportunities. Supporting these ventures is a logical step. By supporting these creations, universities can increase their impact, accelerate students' careers and strengthen the entrepreneurial ecosystem.

%===============================================================
\section{Academia-Driven Entrepreneurship in the Czech Republic and the World}
The commercialisation and success rates of student ventures vary significantly across the world, reflecting differences in ecosystems, funding structures, and cultural attitudes toward entrepreneurship. It is useful to examine the state of innovation in the Czech Republic, to better understand the local potential for academic entrepreneurship.

In a study conducted by the Research, Development and Innovation Council in 2021, the Czech Republic demonstrated strong economic potential and a well-established industrial and research base. Investment in research and development reached a record CZK~121.9 billion (2\% of GDP). The country showed a high publication output, with over 80\% of research results published in indexed journals. Increasing international collaboration has driven excellence in specific scientific fields.\cite{kauffman_tech-transfer}

The study, however, concluded that challenges remain. The PhD completion rates are declining, reflecting low development of researchers' skills and professional capacities. Cooperation between the private and public sectors in research and development is limited. Innovation faces hurdles such as funding shortages and administrative burdens.

The Czech students perceive entrepreneurship as an attractive option for their future career paths. Addressing the systematic weaknesses while capitalising on the existing strengths can improve the Czech Republic’s competitiveness in research, development, and innovation on the global stage.\cite{Antlova_entrepreneurship-cz}

According to infiniteczechgames.com, the gaming industry produces the Czech Republic’s most significant cultural export. Data by the Czech Game Developers Association from 2023 attributed it a turnover of CZK~7.52 billion with more than 98\% of it from abroad.\cite{Žák-gaming, Czech-Developers_INDUSTRY}

Mostly comprised of small studios, game development employed over 2600 workers in 2023---doubling since 2007. 71\% of those workers were under the age of 35 but only 48\% university-educated and only 21.4\% stated formal education as a source of their know-how. These numbers suggest that the field is vibrant and growing, but that the impact of universities and university-provided support is limited.\cite{Žák-gaming, Czech-Developers_INDUSTRY}
